---
title: "hsslda-intro"
author: "meelad amouzgar"
date: "7/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


```{r, echo=TRUE, results = 'hide'}
Tcell_intro = read.csv("~/phd-projects/Rpackage/hsslda/data/TcellHartmann2020_sampleData.txt",sep="\t", stringsAsFactors = T)
```


```{r}
head(Tcell_intro)
colnames(Tcell_intro)
```

You can run hybrid subset selection on a dataset using runHSS(). 
runHSS takes as input a dataframe of cells (rows) and markers/genes (columns), and a vector of your class labels of interest. 
Here we will run HSS using the default, euclidean distance scoring method.
Note that some scoring metrics like silhouette or pixel class entropy (PCE) score will require additional packages.
```{r, echo=TRUE, results = 'hide'}
channels = c('GLUT1', 'HK2', 'GAPDH', 'LDHA', 'MCT1', 'PFKFB4', 'IDH2', 'CyclinB1', 'GLUD12', 'CS', 'OGDH', 'CytC', 'ATP5A', 'S6_p', 'HIF1A')
train.x = Tcell_intro[channels]
train.y = Tcell_intro[['labels']]
hss.result = runHSS(x = train.x, y = train.y, score.method = 'euclidean')

```


The hss.result object contains a few features:

You can visualize the elbow plot, which is ggplot configurable
```{r}
hss.result$ElbowPlot
```

The final output object contains a merged dataframe of your markers, class labels, and newly generated LD axes
```{r}
lda.df = hss.result$`HSS-LDA-result`
head(lda.df)
```

```{r, echo=FALSE}
ggplot2::ggplot(lda.df,ggplot2::aes(x = LD1, y = LD2, color = labels)) +
  ggplot2::geom_point() +
  viridis::scale_color_viridis(discrete = TRUE, option = "viridis") 
```

The final HSS-LDA model is also saved, which you can use to apply the same dimensionality reduction to new data using makeAxes.
```{r, echo=TRUE}
hss.result$`HSS-LDA-model`
```

```{r, echo=FALSE, results = 'hide'}
newdata = train.x
```

```{r, echo=TRUE, results = 'hide'}
lda.df.newData = makeAxes(df = newdata, co =hss.result$`HSS-LDA-model`$scaling)
```





